---
title: "Figure 7 - Nutsch et al Nat Cancer 2024"
output: html_notebook
author: "Charles Tran"
date: "Jul 30, 2024"
---

# Environment Setup
## Load libraries
```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(rstatix)
library(ggpubr)
library(survival)
library(survminer)
library(survivalAnalysis)
library(stringr)
library(gridExtra)
library(svMisc)
```

## Set custom colors for plotting
```{r}
# custom colors for plotting response
response_colors <- c('#d7191c','#fdae61','#abdda4','#2b83ba')
names(response_colors) <- c("CR", "PR", "SD", "PD")
response_colors2 <- c('#fdae61', '#2b83ba')
```

## Define functions (BoxPlot, PFS, OS curves)
```{r}
BoxPlotORR <- function(gene.or.sig.vector, hp) {
  # boxplots for ORR
  ig.plot.list <- list()
  gs <- gene.or.sig.vector
  hp <- hp # horizontal boxplots
  
  if (hp == T) {
    tpm_annotated_scores.temp <- tpm_annotated_scores
    tpm_annotated_scores.temp$ORR_CRPR_SDPD <- factor(tpm_annotated_scores.temp$ORR_CRPR_SDPD, levels = c("SDPD", "CRPR"))
  }
  
  for (v in 1:length(gs)) {
    sig.name <- gs[v]
    sig.name.var <- sym(sig.name)
    
    sig.name.hl <- paste0(sig.name, ".hl", sep="")
    
    # custom expression needed for t_test
    test.expr <- rlang::expr(!! sig.name.var ~ ORR_CRPR_SDPD)
    
    # generate stats for plot
    stat.temp <- tpm_annotated_scores %>% dplyr::select(PATNUM, {{sig.name}}, ORR_CRPR_SDPD) %>% t_test(formula = eval(test.expr)) %>% add_significance("p") %>% add_xy_position(x = "ORR_CRPR_SDPD", fun = "max")
    
    if (hp == T) {
      ig.plot.list[[v]] <- ggplot(tpm_annotated_scores.temp, aes(x=ORR_CRPR_SDPD, y=!!ensym(sig.name))) + 
        geom_boxplot(aes(fill=ORR_CRPR_SDPD), alpha = 0.5, outlier.shape=NA) + 
        geom_point(aes(fill=ORR_CRPR_SDPD), color="black", position = position_jitterdodge(jitter.width=0.1, dodge.width=0.7), pch = 21) + 
        scale_fill_manual(values = rev(response_colors2)) + 
        scale_color_manual(values = rev(response_colors2)) + 
        theme_bw(base_size = 16) +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(hjust = 0.5), legend.position = "none", axis.title.x = element_text(face="italic"), axis.title.y = element_blank()) + 
        labs(x=sig.name) +
        scale_y_continuous(expand=c(0,1)) +
        stat_pvalue_manual(stat.temp, hide.ns=F, tip.length=0, label = "{p}", coord.flip = T) + coord_flip()
      
    } else {
      ig.plot.list[[v]] <- ggplot(tpm_annotated_scores, aes(x=ORR_CRPR_SDPD, y=!!ensym(sig.name))) + 
        geom_boxplot(aes(fill=ORR_CRPR_SDPD), alpha = 0.5, outlier.shape=NA) + 
        geom_point(aes(fill=ORR_CRPR_SDPD), color="black", position = position_jitterdodge(jitter.width=0.1, dodge.width=0.7), pch = 21) + 
        scale_fill_manual(values = response_colors2) + 
        scale_color_manual(values = response_colors2) + 
        theme_bw(base_size = 16) +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(hjust = 0.5), legend.position = "none", axis.title.x = element_blank(), axis.title.y = element_text(face="italic")) + 
        labs(y=sig.name) +
        scale_y_continuous(expand=c(0,1)) +
        stat_pvalue_manual(stat.temp, hide.ns=F, tip.length=0, label = "{p.signif} (p = {p})")
    }
  }
  
  # generate plots in a single grid
  if (hp == T) {
    plot.arranged <- do.call("grid.arrange", c(ig.plot.list, ncol = 1))
  } else {
  plot.arranged <- do.call("grid.arrange", c(ig.plot.list, ncol = length(gs)))
  }
  return(plot.arranged)
}


# PFS
PFSbySig <- function(input.sig) {
  
  grid.draw.ggsurvplot <- function(x){
    survminer:::print.ggsurvplot(x, newpage = FALSE)
  }
  
  PFS.plot.list <- list()
  PFS.survfit.list <- list()
  gs <- input.sig
  
  for (c in 1:length(gs)) {
    sig.name <- gs[c]
    sig.name.var <- paste0("ARM_", sig.name)
    
    PFS.plot.list[[c]] <- ggsurvplot(
      fit = surv_fit(as.formula(paste0("Surv(PFS_MONTHS, PFS_CENSOR) ~ ", sig.name.var)), data = tpm_annotated_scores),
      xlab = "Months", 
      ylab = "PFS",
      legend.title = paste(sig.name, "Score"),
      legend.labs = km.labels, # dynamically generated depending on treatment group(s)
      surv.median.line = "hv",
      palette = "npg",
      risk.table = T,
      pval = T,
      conf.int = F,
      break.time.by = 3,
      risk.table.y.text.col = T,
      tables.theme = theme_cleantable(),
      tables.height = 0.2,
      tables.y.text = FALSE) 
  }
  
  # calculate number of rows needed for plot
  adj.row <- ceiling(length(gs) / 4)
  PFS.all.sig <- arrange_ggsurvplots(PFS.plot.list, print = TRUE, ncol = ifelse(length(gs) < 4, length(gs), 4), nrow = adj.row)
  return(PFS.all.sig)
}

# OS
OSbySig <- function(input.sig) {
  
  grid.draw.ggsurvplot <- function(x){
    survminer:::print.ggsurvplot(x, newpage = FALSE)
  }
  
  OS.plot.list <- list()
  gs <- input.sig
  
  for (d in 1:length(gs)) {
    sig.name <- gs[d]
    sig.name.var <- paste0("ARM_", sig.name)
    
    OS.plot.list[[d]] <- ggsurvplot(
      fit = surv_fit(as.formula(paste0("Surv(OS_MONTHS, OS_CENSOR) ~ ", sig.name.var)), data = tpm_annotated_scores),
      xlab = "Months", 
      xlim = c(0, 36),
      ylab = "OS",
      legend.title = paste(sig.name),
      legend.labs = km.labels,
      surv.median.line = "hv",
      palette = "npg",
      risk.table = T,
      pval = T,
      conf.int = F,
      break.time.by = 3,
      risk.table.y.text.col = T,
      tables.theme = theme_cleantable(),
      tables.height = 0.2,
      tables.y.text = FALSE)
  }
  
  # calculate number of rows needed for plot
  adj.row <- ceiling(length(gs) / 4)
  OS.all.sig <- arrange_ggsurvplots(OS.plot.list, print = TRUE, ncol = ifelse(length(gs) < 4, length(gs), 4), nrow = adj.row)
  return(OS.all.sig)
}
```

# Load data and set analysis options
```{r}
# CITYSCAPE source data: https://ega-archive.org/datasets/EGAD50000000366
tpm_annotated <- readRDS("CITYSCAPE_tpm_annotated.rds")
```

## Load list of gene signatures for the mouse clusters (top 20 genes per cluster from scRNAseq)
```{r}
# csv file can be found in the Supplemental Data section of the online manuscript or is alternatively provided as a RDS file (see "input" folder)
human.genes.from.mouse.sig.intersect <- readRDS("input/human.genes.from.mouse.scrnaseq.clusters.2023-03-20.rds")
```

## Run this to filter only on T+A and removing P+A from analysis (global)
```{r}
# default is analyze all groups
PA.off <- FALSE
km.labels <- c("P+A high", "P+A low", "T+A high", "T+A low")
km.palette <- "npg"
```

# Add signatures scores and other individual gene scores
```{r}
# get metadata layers
tpm_meta <- tpm_annotated[, 1:180]

# add gene signatures from clusters
for (i in 1:length(human.genes.from.mouse.sig.intersect)) {
  gene.sig <- human.genes.from.mouse.sig.intersect[[i]]
  sig.name <- names(human.genes.from.mouse.sig.intersect[i])
  arm.sig <- paste0("ARM_", sig.name, sep="")
  # arm.sig.hl <- paste0(sig.name, ".hl")
  
  if (i == 1) {
    tpm_annotated_scores <- tpm_annotated %>% mutate(!!sym(sig.name) := rowMeans(dplyr::select(., !!!syms(gene.sig))))
  } else {
    tpm_annotated_scores <- tpm_annotated_scores %>% mutate(!!sym(sig.name) := rowMeans(dplyr::select(., !!!syms(gene.sig))))
  }
    # calculate median score and annotate with high or low
    score.median <- tpm_annotated_scores %>% summarise(median = median(!!sym(sig.name))) %>% as.numeric()
    sig.name.hl <- paste0(sig.name, ".hl")
    tpm_annotated_scores <- tpm_annotated_scores %>% mutate(!!sym(sig.name.hl) := ifelse(!!sym(sig.name) >= score.median, "high", "low"))
    
    # add ARM x signature
    tpm_annotated_scores <- tpm_annotated_scores %>% mutate(!!sym(arm.sig) := paste(`ACTARM.2`, !!sym(sig.name.hl), sep = " "))
    
} 

chemokine.ccl5.gene.sets <- list()
chemokine.ccl5.gene.sets[[1]] <- c("CCL5", "CXCR3", "CXCR6", "CCR7")
chemokine.ccl5.gene.sets[[2]] <- c("CCL5", "CXCR3", "CXCR6", "CCR7", "CCR2")
chemokine.ccl5.gene.sets[[3]] <- c("CCL5", "CXCR3", "CXCR6", "CCR7", "CD8A")
chemokine.ccl5.gene.sets[[4]] <- c("CCL5", "CXCR3", "CXCR6", "CCR2")
chemokine.ccl5.gene.sets[[5]] <- c("CCL5", "CXCR3", "CXCR6")
chemokine.ccl5.gene.sets[[6]] <- c("IFNG", "CXCL9", "CD8A", "GZMA", "GZMB", "CXCL10", "PRF1", "TBX21") # add tGE8 score

names(chemokine.ccl5.gene.sets) <- c("CCR7_CXCR3_CXCR6_CCL5", "CCR2_CCR7_CXCR3_CXCR6_CCL5", "CCR7_CXCR3_CXCR6_CCL5_CD8A", "CCR2_CXCR3_CXCR6_CCL5", "CXCR3_CXCR6_CCL5", "tGE8")

# add gene signature scores
for (u in 1:length(chemokine.ccl5.gene.sets)) {
  gene.sig <- chemokine.ccl5.gene.sets[[u]]
  sig.name <- names(chemokine.ccl5.gene.sets[u])
  arm.sig <- paste0("ARM_", sig.name, sep="")
  
  tpm_annotated_scores <- tpm_annotated_scores %>% mutate(!!sym(sig.name) := rowMeans(dplyr::select(., !!!syms(gene.sig))))

  # calculate median score and annotate with high or low
  score.median <- tpm_annotated_scores %>% summarise(median = median(!!sym(sig.name))) %>% as.numeric()
  sig.name.hl <- paste0(sig.name, ".hl")
  tpm_annotated_scores <- tpm_annotated_scores %>% mutate(!!sym(sig.name.hl) := ifelse(!!sym(sig.name) >= score.median, "high", "low"))
    
  # add ARM x signature
  tpm_annotated_scores <- tpm_annotated_scores %>% mutate(!!sym(arm.sig) := paste(`ACTARM.2`, !!sym(sig.name.hl), sep = " "))
} 

# add individual genes
individual.genes <- c("CCL5", "CXCR3", "CXCR6", "CD226", "TIGIT", "CD8A", "CCR7", "CCR2")

for (i in 1:length(individual.genes)) {
  sig.name <- individual.genes[i]
  arm.sig <- paste0("ARM_", sig.name, sep="")
    # calculate median score and annotate with high or low
    score.median <- tpm_annotated_scores %>% summarise(median = median(!!sym(sig.name))) %>% as.numeric()
    sig.name.hl <- paste0(sig.name, ".hl")
    tpm_annotated_scores <- tpm_annotated_scores %>% mutate(!!sym(sig.name.hl) := ifelse(!!sym(sig.name) >= score.median, "high", "low"))
    
    # add ARM x signature
    tpm_annotated_scores <- tpm_annotated_scores %>% mutate(!!sym(arm.sig) := paste(`ACTARM.2`, !!sym(sig.name.hl), sep = " "))
} 

# add ORR
tpm_annotated_scores <- tpm_annotated_scores %>% mutate(ORR_CRPR_SDPD = ifelse(`Response:BCOR` == "CR" | `Response:BCOR` == "PR", "CRPR", "SDPD"))

# add ORR2
tpm_annotated_scores <- tpm_annotated_scores %>% mutate(ORR_CRPRSD_PD = ifelse(`Response:BCOR` == "PD", "PD", "CRPRSD"))

# remove two patients with NA for response data
tpm_annotated_scores <- tpm_annotated_scores %>% filter(!is.na(ORR_CRPR_SDPD))
```

# Plots

## Fig 7a
Plot ORR for each signature comparing CR/PR vs SD/PD
```{r}
plot.list <- list()

for (v in 1:length(human.genes.from.mouse.sig.intersect)) {
  sig.name <- names(human.genes.from.mouse.sig.intersect[v])
  sig.name.var <- sym(sig.name)
  sig.name.hl <- paste0(names(human.genes.from.mouse.sig.intersect[v]), ".hl", sep="")
  
  # custom expression needed for t_test
  test.expr <- rlang::expr(!! sig.name.var ~ ORR_CRPR_SDPD)
  
  # generate stats for plot
  stat.temp <- tpm_annotated_scores %>% dplyr::select(PATNUM, {{sig.name}}, ORR_CRPR_SDPD) %>% t_test(
    formula = eval(test.expr)) %>% add_significance("p") %>% add_xy_position(x = "ORR_CRPR_SDPD", fun = "max")
  
  plot.list[[v]] <- ggplot(tpm_annotated_scores, aes(x=ORR_CRPR_SDPD, y=!!ensym(sig.name))) + 
    geom_boxplot(aes(fill=ORR_CRPR_SDPD), alpha = 0.5, outlier.shape=NA) + 
    geom_point(aes(fill=ORR_CRPR_SDPD), color="black", position = position_jitterdodge(jitter.width=0.1, dodge.width=0.7), pch = 21) + 
    scale_fill_manual(values = response_colors2) + 
    scale_color_manual(values = response_colors2) + 
    theme_bw(base_size = 16) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(hjust = 0.5), 
          legend.position = "none", axis.title.x = element_blank()) + 
    labs(y=paste(sig.name, "Score")) +
    scale_y_continuous(expand=c(0,1)) +
    stat_pvalue_manual(stat.temp, hide.ns=T, tip.length=0, label = "{p.signif} (p = {p})") 
}

# generate plots in a single grid
plot.arranged <- do.call("grid.arrange", c(plot.list, ncol = 6))
print(plot.arranged)
```


## Fig 7b
Analysis of CCL5, CXCR3, CXCR6, CD8A seprately and association with ORR 
```{r}
# boxplots for ORR
ig.plot.list <- list()

individual.genes <- c("CCL5", "CXCR3", "CXCR6", "CD8A")

for (v in 1:length(individual.genes)) {
  sig.name <- individual.genes[v]
  sig.name.var <- sym(sig.name)
  sig.name.hl <- paste0(sig.name, ".hl", sep="")
  
 # custom expression needed for t_test
  test.expr <- rlang::expr(!! sig.name.var ~ ORR_CRPR_SDPD)
  
 # generate stats for plot
  stat.temp <- tpm_annotated_scores %>% dplyr::select(PATNUM, {{sig.name}}, ORR_CRPR_SDPD) %>% t_test(formula = eval(test.expr)) %>% add_significance("p") %>% add_xy_position(x = "ORR_CRPR_SDPD", fun = "max")
  
  ig.plot.list[[v]] <- ggplot(tpm_annotated_scores, aes(x=ORR_CRPR_SDPD, y=!!ensym(sig.name))) + 
    geom_boxplot(aes(fill=ORR_CRPR_SDPD), alpha = 0.5, outlier.shape=NA) + 
    geom_point(aes(fill=ORR_CRPR_SDPD), color="black", position = position_jitterdodge(jitter.width=0.1, dodge.width=0.7), pch = 21) + 
    scale_fill_manual(values = response_colors2) + 
    scale_color_manual(values = response_colors2) + 
    theme_bw(base_size = 16) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(hjust = 0.5), legend.position = "none", axis.title.x = element_blank(), axis.title.y = element_text(face="italic")) + 
    labs(y=sig.name) +
    scale_y_continuous(expand=c(0,1)) +
    stat_pvalue_manual(stat.temp, hide.ns=F, tip.length=0, label = "{p.signif} (p = {p})") 
}

# horizontal boxplot version of above, only p value (no asterix or 'ns')
tpm_annotated_scores$ORR_CRPR_SDPD <- factor(tpm_annotated_scores$ORR_CRPR_SDPD, levels = c("SDPD", "CRPR"))

for (v in 1:length(individual.genes)) {
  sig.name <- individual.genes[v]
  sig.name.var <- sym(sig.name)
  sig.name.hl <- paste0(sig.name, ".hl", sep="")
  
 # custom expression needed for t_test
  test.expr <- rlang::expr(!! sig.name.var ~ ORR_CRPR_SDPD)
  
   # generate stats for plot
  stat.temp <- tpm_annotated_scores %>% dplyr::select(PATNUM, {{sig.name}}, ORR_CRPR_SDPD) %>% t_test(formula = eval(test.expr)) %>% add_significance("p") %>% add_xy_position(x = "ORR_CRPR_SDPD", fun = "max")
  
  ig.plot.list[[v]] <- ggplot(tpm_annotated_scores, aes(x=ORR_CRPR_SDPD, y=!!ensym(sig.name))) + 
    geom_boxplot(aes(fill=ORR_CRPR_SDPD), alpha = 0.5, outlier.shape=NA) + 
    geom_point(aes(fill=ORR_CRPR_SDPD), color="black", position = position_jitterdodge(jitter.width=0.1, dodge.width=0.7), pch = 21) + 
    scale_fill_manual(values = rev(response_colors2)) + 
    scale_color_manual(values = rev(response_colors2)) + 
    theme_bw(base_size = 16) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(hjust = 0.5), legend.position = "none", axis.title.x = element_text(face="italic"), axis.title.y = element_blank()) + 
    labs(y=sig.name) +
    scale_y_continuous(expand=c(0,1)) +
    stat_pvalue_manual(stat.temp, hide.ns=F, tip.length=0, label = "{p}", coord.flip=TRUE) + coord_flip() 
}

# generate plots in a single grid
plot.arranged <- do.call("grid.arrange", c(ig.plot.list, ncol = 1)) # for horizontal version (1 column)

print(plot.arranged)
```


## Fig 7c
Forest plots for CCL5, CXCR3, CXCR6, CD8A
```{r}
# generate forest plots for OS HR
OS.HR.ig.high.list <- list()
OS.HR.ig.low.list <- list()

individual.genes.subset <- c("CCL5", "CXCR3", "CXCR6", "CD8A")

for (n in 1:length(individual.genes.subset)) {
  sig.name <- individual.genes.subset[n]
  sig.name.hl <- paste0(sig.name, ".hl", sep="")
  
  tpm_annotated_scores_high <- tpm_annotated_scores %>% filter(!!sym(sig.name.hl) == "high")
  tpm_annotated_scores_low <- tpm_annotated_scores %>% filter(!!sym(sig.name.hl) == "low")
  
  # factor to compare tira_atezo vs placebo_atezo
  tpm_annotated_scores_high$ACTARM.2 <- factor(tpm_annotated_scores_high$ACTARM.2, levels = c( "placebo_atezo", "tira_atezo"))
  tpm_annotated_scores_low$ACTARM.2 <- factor(tpm_annotated_scores_low$ACTARM.2, levels = c("placebo_atezo", "tira_atezo"))

  # calculate coxph
  OS.HR.ig.sig.high <- coxph( Surv(OS_MONTHS, OS_CENSOR) ~ ACTARM.2, data = tpm_annotated_scores_high)
  OS.HR.ig.sig.low <- coxph( Surv(OS_MONTHS, OS_CENSOR) ~ ACTARM.2, data = tpm_annotated_scores_low)
  
  # convert to df
  OS.HR.ig.sig.high.df <- cox_as_data_frame(OS.HR.ig.sig.high) %>% mutate(gene.sig = paste(sig.name, "high"))
  OS.HR.ig.sig.low.df <- cox_as_data_frame(OS.HR.ig.sig.low) %>% mutate(gene.sig = paste(sig.name, "low"))
  
  # add to list of dataframes
  OS.HR.ig.high.list[[n]] <- OS.HR.ig.sig.high.df
  OS.HR.ig.low.list[[n]] <- OS.HR.ig.sig.low.df
}

# combine df
OS.HR.ig.high.sig.df <- Reduce(full_join, OS.HR.ig.high.list)
OS.HR.ig.low.sig.df <- Reduce(full_join, OS.HR.ig.low.list)

OS.HR.ig.all.sig.df <- rbind(OS.HR.ig.high.sig.df, OS.HR.ig.low.sig.df)

highlow <- c("low", "high")
named.ig.sig <- NULL
for (m in 1:length(individual.genes.subset)) {
   sig.name <- individual.genes.subset[m]
   for (r in 1:length(highlow)) {
     temp.name <- paste(sig.name, highlow[r])
     named.ig.sig <- c(named.ig.sig, temp.name)
   }
}

# re-order levels
OS.HR.ig.all.sig.df$gene.sig = factor(OS.HR.ig.all.sig.df$gene.sig, levels = rev(named.ig.sig))

# plot HR forest plot (with p values)
ggplot(OS.HR.ig.all.sig.df, aes(x = gene.sig, y = HR, fill=gene.sig)) + geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width=0.2, position=position_dodge(0.8)) + geom_point(aes(fill=gene.sig), pch=21, color="black", size=6) + geom_hline(yintercept=1, linetype="dashed", color="red", linewidth = 1) + theme_bw(base_size=18) + coord_flip(clip = "off") + scale_y_continuous(breaks = seq(0.1, 1.9, by = 0.2)) + theme(panel.grid=element_blank(), legend.position = "none", plot.margin = unit(c(1, 8, 1, 1), "lines")) + xlab("Gene") + 
geom_text(aes(label = paste(sprintf('%.3f', p), " | ", sprintf('%.2f', HR), sep="")), y = 2, check_overlap = TRUE, hjust = 1)
```


## Fig 7d
KM plots for individual genes CD8A, CCL5, CXCR3, CXCR6 by treatment group (T+A, P+A)
```{r}
# generate OS plots
OS.ig.plot.list <- list()
OS.ig.survfit.list <- list()

individual.genes.subset <- c("CD8A", "CXCR3", "CXCR6", "CCL5")

for (e in 1:length(individual.genes.subset)) {
  sig.name <- individual.genes.subset[e]
  sig.name.var <- paste0("ARM_", sig.name)
  
  OS.ig.plot.list[[e]] <- ggsurvplot(
    fit = survfit(as.formula(paste0("Surv(OS_MONTHS, OS_CENSOR) ~ ", sig.name.var)), data = tpm_annotated_scores),
    xlab = "Months", 
    xlim = c(0, 36),
    ylab = "OS",
    legend.title = paste(sig.name),
   legend.labs = km.labels,
    surv.median.line = "hv",
    palette = "npg",
    risk.table = T,
    pval = T,
    conf.int = F,
    break.time.by = 3,
    risk.table.y.text.col = T,
    tables.theme = theme_cleantable(),
    tables.height = 0.2,
    tables.y.text = FALSE)
}

OS.ig.all.sig <- arrange_ggsurvplots(OS.ig.plot.list, print = TRUE, ncol = 2, nrow = 2)
print(OS.ig.all.sig)

```


## Fig 7e
Generate ORR boxplots for CXCR3+CXCR6+CCL5 signature
```{r}
sig.temp.gs <- c("CXCR3_CXCR6_CCL5")

temp.arranged <- BoxPlotORR(sig.temp.gs, hp = T)
print(temp.arranged)
```


## Fig 7f
Forest plot for CXCR3+CXCR6+CCL5 signature and OS HR
```{r}
OS.HR.high.list <- list()
OS.HR.low.list <- list()

# for purposes of paper, only plot selected sig.temp.gs signatures
sig.temp.gs <- c("CXCR3_CXCR6_CCL5")

for (k in 1:length(sig.temp.gs)) {
  # gene.sig <- human.genes.from.mouse.sig.intersect[[k]]
  sig.name <- sig.temp.gs[k]
  sig.name.hl <- paste0(sig.temp.gs[k], ".hl", sep="")
  
  tpm_annotated_scores_high <- tpm_annotated_scores %>% filter(!!sym(sig.name.hl) == "high")
  tpm_annotated_scores_low <- tpm_annotated_scores %>% filter(!!sym(sig.name.hl) == "low")
  
  # factor to compare tira_atezo vs placebo_atezo
  tpm_annotated_scores_high$ACTARM.2 <- factor(tpm_annotated_scores_high$ACTARM.2, levels = c( "placebo_atezo", "tira_atezo"))
  tpm_annotated_scores_low$ACTARM.2 <- factor(tpm_annotated_scores_low$ACTARM.2, levels = c("placebo_atezo", "tira_atezo"))

  # calculate coxph
  OS.HR.sig.high <- coxph( Surv(OS_MONTHS, OS_CENSOR) ~ ACTARM.2, data = tpm_annotated_scores_high)
  OS.HR.sig.low <- coxph( Surv(OS_MONTHS, OS_CENSOR) ~ ACTARM.2, data = tpm_annotated_scores_low)
  
  # convert to df
  OS.HR.sig.high.df <- cox_as_data_frame(OS.HR.sig.high) %>% mutate(gene.sig = paste(sig.name, "high"))
  OS.HR.sig.low.df <- cox_as_data_frame(OS.HR.sig.low) %>% mutate(gene.sig = paste(sig.name, "low"))
  
  # add to list of dataframes
  OS.HR.high.list[[k]] <- OS.HR.sig.high.df
  OS.HR.low.list[[k]] <- OS.HR.sig.low.df

}

# combine df
OS.HR.high.sig.df <- Reduce(full_join, OS.HR.high.list)
OS.HR.low.sig.df <- Reduce(full_join, OS.HR.low.list)

OS.HR.all.sig.df <- rbind(OS.HR.high.sig.df, OS.HR.low.sig.df)

highlow <- c("low", "high")
named.sig <- NULL
for (n in 1:length(sig.temp.gs)) {
   sig.name <- sig.temp.gs[n]
   for (l in 1:length(highlow)) {
     temp.name <- paste(sig.name, highlow[l])
     named.sig <- c(named.sig, temp.name)
   }
}

# re-order levels
OS.HR.all.sig.df$gene.sig = factor(OS.HR.all.sig.df$gene.sig, levels = rev(named.sig))

ggplot(OS.HR.all.sig.df, aes(x = gene.sig, y = HR, fill=gene.sig)) + geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width=0.2, position=position_dodge(0.8)) + geom_point(aes(fill=gene.sig), pch=21, color="black", size=6) + geom_hline(yintercept=1, linetype="dashed", color="red", size = 1) + theme_bw(base_size=18) + coord_flip(clip = "off") + scale_y_continuous(breaks = seq(0.1, 1.9, by = 0.2)) + theme(panel.grid=element_blank(), legend.position = "none", plot.margin = unit(c(1, 8, 1, 1), "lines")) + xlab("Gene") + 
geom_text(aes(label = paste(sprintf('%.3f', p), " | ", sprintf('%.2f', HR), sep="")), y = 2, check_overlap = TRUE, hjust = 1)
```


## Fig 7g
KM plot for CXCR3+CXCR6+CCL5 signature for T+A and P+A groups
```{r}
OS.surv.temp <- OSbySig(sig.temp.gs)
print(OS.surv.temp)
```
